Towards Bringing Together Numerical Methods for Partial Differential Equation and Image Processing

### 2020-05-07 paper/paper.tex
Correcting the paper

Classical methods for performing such simulations are based on partial differential equations (PDEs) solvers. The simulation setup is first formalized as a mathematical model in the form of a time-dependent differential equation. In itself, this equation is then transformed and brought into a suitable for solving form. A common technique is the finite difference method (FDM)~\cite{bashar2011,petersson2018,zhang2014,nahit2011}. This method can provide accurate results at the cost of large computational time. The generated results are in the form of raw numbers representing the velocities and pressure fields which have to be visualized separately. Our method aims to generate straight the visualizations while needing much lower computational time.

@@

Lastly, we briefly want to motivate why we propose exactly this approach. Images represent a well-defined input space as every pixel can take a limited range of values --- $[0, 255]$. This makes the handling of data in image form convenient. Based on the previous works, we assume that fully convolutional neural network can process information in this form more easily. Furthermore, convolutional neural networks are well-established models for performing image processing tasks. We were interested in how a numerical task can be turned into an image processing task as in certain situations, the image results have the priority. We also think that the image processing can be done much quicker in a more optimal way. Those considerations urged us to conduct the research presented in this paper.

@@

In CFD, DNNs have been used in a wide variety of contexts. Turbulence models have been developed in~\cite{beck2018},~\cite{pfeiffer2019} has developed a network that predicts the motion of organs during operation and~\cite{tompson2016} uses a highly tailored neural network to solve Euler equations for fluid flow and thus accelerate the simulation process~\cite{thuerey2018} solves the problem of the Reynolds-Averaged Navier-Stokes (RANS) simulation similarly to us. They do not use cGANs, do not target the image representation of the simulation and tackle different equation from us but they too try to map initial conditions to a solution through an end-to-end model.

@@

The choice of concrete values for the parameters is deliberate. All of the values are chosen so that the Reynolds number~\cite{trinh2010critical} of the simulations in the range of $[90, 450]$. This keeps the flow laminar while still making it interesting enough. We were interested whether the built models can predict the emerging \emph{Kármán vortex street}~\cite{karaman1911} behind the object in the channel. Thus the Reynolds numbers were chosen so that the effect can occur.

@@

\subsection{Training approach and networks details}
We base our generative models almost entirely on~\cite{pix2pix}. We use the conditional GAN approach to train a generator network that can perform image-to-image translation. As explained in~\cite{pix2pix}, the traditional GAN method uses a random vector $z$ as in input to the generator network $G$ to generate output $y$, $G:z \rightarrow y$. Conditional GANs also feed an input image $x$ to the generator, $G: x,z \rightarrow y$.~\cite{pix2pix} and~\cite{wang2016} suggest that in certain cases the usage of $z$ can be usefully but we decided not to include for our generator as we want a deterministic network.

We adopt the objective for the discriminator network as we modify it slightly by leaving out the random vector $z$.
\begin{equation}
\mathcal{L}_{cGAN} (G, D) = (\mathbb{E}[\log D (x,y)] + \mathbb{E}[\log D (x, G(x))])/2
\end{equation}
where $x$ is the input image and $y$ is the target image. In contrast to unconditional GANs, both the generator and the discriminator network have access to the input image. The objective is divided by two to slow down the training of the discriminator relative to the generator as suggested by~\cite{pix2pix}.

The objective for the generator network is comprised of two parts --- the value of the discriminator as well as a $L1$ distance loss between the target and the predicted images. According to~\cite{pix2pix} the $L1$ loss promotes less blurring and captures the low frequency details of the images. The $L1$ loss is given by:
\begin{equation}
\mathcal{L}_{L1} (G) = \mathbb{E}[\left\lVert y - G(x)\right\rVert_1]
\end{equation}
The final object for the generator is thus:
\begin{equation}
G^* = \argminA_G \max_D \mathcal{L}_{cGAN}  + \lambda \mathcal{L}_{L1} (G)
\end{equation}
For all of the models we used $\lambda = 100$.

@@

For the discriminator, we follow the method of pix2pix and we use their PatchGAN.\@ This is a convolutional network that examines only patches of the input. It tries to guess if each patch is from a real or generated image. We use patches of size $256\times 256$ pixels.

\noindent\textbf{Training details:} We train the described three models with the generated datasets. When loading the images in memory, we first resize them to $1024\times 256$ (height $\times$ width). We then apply random crops as well as add random noise to each channel of the images. We do this to force the generator to learn the actual features of the simulation and make over-fitting harder. Before the images are fed into the networks, we also multiply them with a mask of the object in the channel. This is an image that has zero value in the area where the object is located and values of one for every other location. The multiplication with the mask results in an input image with values of zero in the area of the object. The object mask itself is also given as an input to the generator network.

@@

For the training procedure, we follow the standard approach proposed in~\cite{goodfellow2014}. With each mini-batch, we first optimize the discriminator and then the generator with the discussed objectives. We use Stochastic Gradient Descent~\cite{kiefer1952} with the Adam optimizer~\cite{kingma2014} with a learning rate of $0.0002$ and momentum parameters $\beta_1=0.9$ and $\beta_2=0.999$. The used batch size for the constant and inflow speed models was set to 3 and for the viscosity-density models to 4. Those are relatively small numbers for batch sizes but~\cite{pix2pix} suggests that the U-Net architecture benefits from small batches in image-to-image translation problems. Our experiments confirmed this.

@@

On one hand, we are dealing with image data that is meant to be perceived by a human observer. This implies that full accuracy is not essential and therefore we place the importance on the perceived qualities of the predicted frames. This makes sense as in our case a human is concerned with the general development of the simulation over time. Under those considerations we have chosen appropriate metrics to compare the predicted simulation images with the real ones. In this aspect, the first of the metrics is \emph{Peak Signal to Noise Ratio} (PSNR). It is related to the Mean Squared Error (MSE) and it is given by
\begin{equation}
PSNR = 10 \cdot \log_{10}\left(\frac{255^2}{MSE}\right)
\end{equation}
PSNR is measured in decibel (dB) and it is a common metric used to evaluate lossy compression algorithms~\cite{psnr1,psnr2}. It measures the difference between the original and reconstructed images and considers the visual fidelity of the reconstruction. This makes the metric suitable in our case. Typical PSNR values of lossy image and video compression algorithms are between $30$ and $50$ dB.

When considering the perceived image quality, we also have decided to use correlation to compare the predicted and the real simulation frames. We think that if the predicted images differ only in a factor from the real ones, the models can accurately capture the relative difference between the pixels of the target image. For this reason, we opted for also evaluating the results through statistical correlation.

On the other hand, as we are dealing with a numerical task, we also wanted to evaluate the results more objectively. Classical solvers are evaluated in terms of the actual differences between the real and generated result and not the perceived differences. For those reasons, we have chosen two more metrics --- average absolute difference and maximal absolute differences. We calculate them based on differences between the corresponding pixels of the predicted and the real simulation frames. Both metrics are then averaged across the test set and are given in percentual values. The evaluation with these metrics should represent how well our models perform the simulation in the context of a CFD system.

\noindent\textbf{Details Evaluation:}
\emph{In this subsection, we want to give several details on the training and evaluating approach.}
\begin{itemize}
\item[$\cdot$] All models were trained from scratch and no pretraining or transfer-learning was undertaken.
\item[$\cdot$] For each model, the whole training and evaluating procedure was done ten times with different seeds for the random number generators of our python scripts. We report the results of each run and give them in a form of error plots or explicitly give the standard deviation of all runs.
\item[$\cdot$] All of the reported results for the individual runs are averaged across the simulation frames in the training sets.
\item[$\cdot$] While performing the experiments, we did not tweak the hyperparameters of the models excessively. We did, however, tried several sets of hyperparameters until the models achieved satisfactory results. The exact hyperparameters are given in \refapp{app1}.
\end{itemize}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{lcc|cc}
      \textbf{Model type} & \multicolumn{2}{c|}{\textbf{Max difference}}  &  \multicolumn{2}{c}{\textbf{Average difference}}\\
      \hline 
      \multicolumn{3}{c|}{}&&\\
                 & \underline{With pr.} & \underline{without pr.}  &  \underline{With pr}.& \underline{without pr.}\\

@@

\noindent\textbf{Performance:} All of the models were evaluated on an Nvidia GTX 980Ti GPU.\@In all cases, the forward propagation of a single simulation frame took between 5 and 20 ms (milliseconds). In comparison, during the data generation, HiFlow needed between 1000 and 1500 ms for a single simulation timestep while using 12 MPI processes. This does not include the time needed for rendering the results of the simulation. To note is that this is not a one-to-one comparison as HiFlow runs on the CPU and uses MPI and OpenMP for parallelization. We, however, think that the results show the potential of our approach when it comes to performance.

@@

This configuration results in a receptive field of $256\times256$. This means that for each $256\times256$ region of the input, the discriminator tries to guess if the region is from a real image or from a generated one.

>>>>

Classical methods for performing such simulations are based on partial differential equations (PDEs) solvers. The simulation setup is first formalized as a mathematical model in the form of a time-dependent differential equation. In itself, this equation is then transformed and brought into a suitable for solving form. In a general case, basic technique that can be used is the finite difference method (FDM)~\cite{bashar2011,petersson2018,zhang2014,nahit2011}. This method can provide accurate results at the cost of large computational time. The generated results are in the form of raw numbers representing the velocities and pressure fields which have to be visualized separately. Our method aims to generate straight the visualizations while needing much lower computational time.

@@

Lastly, we briefly want to motivate why we propose exactly this approach. Images represent a well-defined input space as every pixel can take a limited range of values --- $[0, 255]$. This makes the handling of data in image form convenient. Based on the previous works, we assume that fully convolutional neural network can process information in this form more easily. Furthermore, convolutional neural networks are well-established models for performing image processing tasks. We were interested in how a numerical task can be turned into an image processing task as in certain situations, the image results have the priority.
We believe that this image processing approach can be used to generate a simulation image much faster than a numerical approach with some loss in accuracy. The accuracy loss is the subject of our research.

@@

In CFD, DNNs have been used in a wide variety of contexts. Turbulence models have been developed in~\cite{beck2018},~\cite{pfeiffer2019} has developed a network that predicts the motion of organs during operation and~\cite{tompson2016} uses a highly tailored neural network to solve Euler equations for fluid flow and thus accelerate the simulation process~\cite{thuerey2018} solves the problem of the Reynolds-Averaged Navier-Stokes (RANS) simulation similarly to us. They do not use cGANs, do not target the image representation of the simulation and tackle different equation from us but they too try to map initial conditions to a solution through an end-to-end model. Our approach is not exactly comparable to the one of~\cite{thuerey2018} as we are considering a different numeric problem (the Navier-Stokes equations). Also,~\cite{thuerey2018} tries to map the boundary conditions of the problem to the solutions. We, on the other hand, map a solution of the studied simulation to the next one.

@@

The choice of concrete values for the parameters is deliberate. All of the values are chosen so that the Reynolds number~\cite{trinh2010critical} of the simulations in the range of $[90, 450]$. This causes a transitional flow with vortex shedding effect. We were interested whether the built models can predict the emerging \emph{Kármán vortex street}~\cite{karaman1911} behind the object in the channel. Thus the Reynolds numbers were chosen so that the effect can occur.

@@

\subsection{Training approach and networks details}
We base our generative models almost entirely on~\cite{pix2pix}. We use the conditional GAN approach to train a generator network that can perform image-to-image translation. As explained in~\cite{pix2pix}, the traditional GAN method uses a random vector $z$ as in input to the generator network $G$ to generate output $y$, $G:z \rightarrow y$. Conditional GANs also feed an input image $x$ to the generator, $G: x,z \rightarrow y$.~\cite{pix2pix} and~\cite{wang2016} suggest that in certain cases the usage of $z$ can be usefully but we decided not to include for our generator as we want a deterministic network. The discriminator network is modeled with the function $D:x,y \rightarrow v$ that evaluates the likelihood of $y$ being a real image. To note is that the discriminator network has access to the real image $x$ and tries to guess, if $y$ is the real or generated output.

We adopt the objective for the discriminator network and we modify it slightly by leaving out the random vector $z$.
\begin{equation}
  \mathcal{L}_{cGAN} (G, D) = (\mathbb{E}[\log D (x,y)] + \mathbb{E}[\log D (x, G(x))])/2 \\
  = (\log D (x,y) + \log D (x, G(x)))/2
\end{equation}
where $x$ is the input image and $y$ is the target image. We leave out the expected value calculation as we do not use the random vector $z$ in our objective. In contrast to unconditional GANs, both the generator and the discriminator network have access to the input image. The objective is divided by two to slow down the training of the discriminator relative to the generator as suggested by~\cite{pix2pix}.

The objective for the generator network is comprised of two parts --- the value of the discriminator as well as a $L1$ distance loss between the target and the predicted images. According to~\cite{pix2pix} the $L1$ loss promotes less blurring and captures the low frequency details of the images. The $L1$ loss is given by:
\begin{equation}
\mathcal{L}_{L1} (G) = \mathbb{E}[\left\lVert y - G(x)\right\rVert_1]
\end{equation}
The final object for the generator is thus:
\begin{equation}
G^* = \argminA_G \max_D \mathcal{L}_{cGAN}  + \lambda \mathcal{L}_{L1} (G)
\end{equation}
For all of the models we used $\lambda = 100$ as done in~\cite{pix2pix}. We also experimented with lower values for $\lambda$ but results were not as good.

@@

For the discriminator, we follow the method of~\cite{pix2pix} and we use their PatchGAN discriminator network.\@ This is a convolutional network that classifies patches of the input as real or predicted. To not is that the whole image is given as an input. The majority of the results in~\cite{pix2pix} show that a patches of size $70\times 70$ yield the best results but in our case, the experiments showed otherwise. We, therefore, we opted out for using patches of size $286\times 286$ pixels.

\noindent\textbf{Training details:} We train the described three models with the generated datasets. When loading the images in memory, we first resize them to an appropriated for a network size of $1024\times 256$ (width $\times$ height). We then apply random crops as well as add random noise to each channel of the images. We do this to force the generator to learn the actual features of the simulation and make over-fitting harder. Before the images are fed into the networks, we also multiply them with a mask of the object in the channel. This is an image that has zero value in the area where the object is located and values of one for every other location. The multiplication with the mask results in an input image with values of zero in the area of the object. The predicted images are also multiplied with the mask. This procedure lets us examine how well the network generates the pixels in the places that we are interested in --- the places where the flow occurs. The object mask itself is also given as an input to the generator network. This gives the network the information of where the object is. This information cannot be embedded in the other input images as the zeroed values in the location of the object are interpreted as velocity values by network.

@@

For the training procedure, we follow the standard approach proposed in~\cite{goodfellow2014}. With each mini-batch, we first optimize the discriminator and then the generator with the discussed objectives. We use Stochastic Gradient Descent~\cite{kiefer1952} with the Adam optimizer~\cite{kingma2014} with a learning rate of $0.0002$ and standard momentum parameters $\beta_1=0.9$ and $\beta_2=0.999$. The used batch size for the constant and inflow speed models was set to $3$ and for the viscosity-density models to $4$. Those are relatively small numbers for batch sizes but~\cite{pix2pix} suggests that the U-Net architecture benefits from small batches in image-to-image translation problems. Our experiments confirmed this.

@@

On one hand, we are dealing with image data that is meant to be perceived by a human observer. This implies that full accuracy is not essential and therefore we place the importance on the perceived qualities of the predicted frames. This makes sense as in our case a human is concerned with the general development of the simulation over time. Under those considerations we have chosen appropriate metrics to compare the predicted simulation images with the real ones. In this aspect, the first of the metrics is \emph{Peak Signal to Noise Ratio} (PSNR). It is related to the Mean Squared Error (MSE) and it is given by
\begin{equation}
PSNR = 10 \cdot \log_{10}\left(\frac{255^2}{MSE}\right)
\end{equation}
PSNR is measured in decibel (dB) and it is a common metric used to evaluate lossy compression algorithms~\cite{psnr1,psnr2}. It measures the difference between the original and reconstructed images and considers the visual fidelity of the reconstruction. This makes the metric suitable in our case. Typical PSNR values of lossy image and video compression algorithms are between $30$ and $50$ dB. Higher PSNR values are desired, as the value of PSNR is inversely proportional to the value of MSE.

When considering the perceived image quality, we also have decided to use correlation to compare the predicted and the real simulation frames. We think that if the predicted images differ only in a factor from the real ones, the models can accurately capture the relative difference between the pixels of the target image. For this reason, we opted for also evaluating the results through statistical correlation between the predicted and the real images. Naturally, we want to see if the models can generated images with correlation close to $1.0$.

On the other hand, as we are dealing with a numerical task, we also wanted to evaluate the results more objectively. Classical solvers are evaluated in terms of the actual differences between the real and generated result and not the perceived differences. For those reasons, we have chosen two more metrics --- average absolute difference and maximal absolute differences. We calculate them based on differences between the corresponding pixels of the predicted and the real simulation frames. Both metrics are then averaged across the test set and are given in percentual values. The evaluation with these metrics should represent how well our models perform the simulation in the context of a CFD system. As their are both difference measures, lower values are desired for these metrics.

\noindent\textbf{Evaluation Details:}
\emph{In this subsection, we want to give several details on the training and evaluating approach.}
\begin{itemize}
\item[$\cdot$] All models were trained from scratch and no pretraining or transfer-learning was undertaken.
\item[$\cdot$] For each model, the whole training and evaluating procedure was done ten times with different seeds for the random number generators of our python scripts. We report the results of each run and give them in a form of error plots or explicitly give the standard deviation of all runs.
\item[$\cdot$] All of the reported results for the individual runs are averaged across the simulation frames in the training sets.
\item[$\cdot$] While performing the experiments, we did not tweak the hyperparameters of the models excessively. We did, however, tried several sets of hyperparameters until the models achieved satisfactory results. The exact hyperparameters are given in \refapp{app1}.
\end{itemize}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{lcc|cc}
      \textbf{Model type} & \multicolumn{2}{c|}{\textbf{Max difference(\%) }}  &  \multicolumn{2}{c}{\textbf{Average difference(\%)}}\\
      \hline 
      \multicolumn{3}{c|}{}&&\\
                 & \underline{With pr.} & \underline{without pr.}  &  \underline{With pr}.& \underline{without pr.}\\

@@

\noindent\textbf{Performance:} All of the models were evaluated on an Nvidia GTX 980Ti GPU.\@In all cases, the forward propagation of a single simulation frame took between 5 and 20 ms (milliseconds). In comparison, during the data generation, HiFlow needed between 1000 and 1500 ms for a single simulation timestep while using 12 MPI processes. This does not include the time needed for rendering the results of the simulation. To note is that this is not a one-to-one comparison as HiFlow runs on the CPU and uses MPI and OpenMP for parallelization. Also, we performed all the real simulations with HiFlow on a Intel Xeon E5-2650 v4 CPU.

We, however, think that the results show the potential of our approach when it comes to performance.

@@

This configuration results in a receptive field of $286\times286$. This means that for each $256\times256$ region of the input, the discriminator tries to guess if the region is from a real image or from a generated one.
